Efficiently Handling Guards in the Low Level Design of RPython's Tracing JIT

### 2012-08-10 talk/vmil2012/paper.tex
Write some conclusion-ish words
rewrite conclusion/related work a bit

\section{Conclusion}
\label{sec:Conclusion}

\todo{conclusion}

>>>>

\section{Conclusion}
\label{sec:Conclusion}
In this paper we have concentrated on guards, an operation typically found in
tracing just-in-time compilers and used to denote points of possible control
flow divergence in recorded traces.
We described how, based on the observation that guards are a frequent operation
in traces and that they do not fail often, guards have been implemented in the
high and low level components of RPython's tracing JIT compiler.

Finally we have presented experimental data collected using the standard PyPy
benchmark set to evaluate previous observations and assumptions. Our
experiments showed that, as previously assumed, guards are a very common
operation in traces. At the same time guards are associated with a high
overhead, because for all compiled guards information needs to be
stored to restore the execution state in case of a bail-out. The measurements
showed that the compression techniques used in PyPy effectively reduce the
overhead of guards, while it still produces a significant overhead. The results
also showed that guard failure is a local event: there are few
guards that fail at all, and even fewer that fail very often.
These numbers validate the design decision of reducing the overhead of
successful guard checks as much as possible while paying a higher price in the
case of bailout due to having to decode compressed state representation.
The compressed state representation is reduces the memory footprint of rarely
used data.

Based on the observation that most guards do not fail very often or at all it
would be worth exploring if a more aggressive compression scheme for guards
would be worth the memory saving in contrast to the increased decoding
overhead. Based on the same observation we would like to explore the concept of
LuaJIT's sparse snapshots and its applicability to PyPy.
